---
title: Homework 1
author: Erin Yang
format: pdf
---

- introduce county, basically manhattan. 
- people of color in certain neighborhoods, also overlaps w low income
- introduce diesel particulate matter - pg 16 on pdf
- describe health effects
- analysis: if u compare the two maps the red parts are like in teh same parts of the county
- propose some explanations for why there is so much overlap
- list equity issues with said overlap
- restrictions: diesel pm was actually calculated at the census tract level, not block



```{r}
library(tidyverse)
library(sf)
library(here)
library(tmap)

# read in geodatabase of EJScreen data at the Census Block Group level
ejscreen <- sf::st_read(here::here("data", "ejscreen","EJSCREEN_2023_BG_StatePct_with_AS_CNMI_GU_VI.gdb")) 

# filter to a state you are interested in
state <- ejscreen %>%
  dplyr::filter(ST_ABBREV == "NY") 

# filter to a county you are interested in
county <- ejscreen %>%
  dplyr::filter(CNTY_NAME %in% c("New York County"))

# find the average values for all variables within counties
#state_by_county <- aggregate(state, by = list(county$CNTY_NAME), FUN = mean)
```


```{r}
tm_shape(county) +
    tm_polygons(
        fill = 'PEOPCOLORPCT',
        fill.scale = tm_scale(values = c('green', 'red')),
        fill.legend = tm_legend(title = 'Percentage People of Color'),
        fill_alpha = 0.6
    ) +
    tm_scalebar() +
    tm_title('Percentage People of Color by Census Block in New York County', size = 0.5)
```


```{r}
tm_shape(county) +
    tm_polygons(
        fill = 'DSLPM',
        fill.scale = tm_scale(values = c('green', 'red')),
        fill.legend = tm_legend(title = 'Percentile Diesel Particulate Matter (μg/m3)'),
        fill_alpha = 0.6
    ) +
    tm_scalebar() +
    tm_title('Percentile Diesel Particulate Matter by Census Block in New York County', size = 0.5)
```

